Categorical Predictor Embeddings and Attention Based Neural Networks in Actuarial Modeling

### 2021-04-07 manuscript/manuscript.Rmd
Changed models to reflect axes
Update manuscript section 6

---
title: Categorical Predictor Embeddings in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
header-includes: |
    \usepackage{amsmath}
    \usepackage{mathtools}
    \usepackage{bm}
    \usepackage{bbm}
    \usepackage{float}
    \DeclarePairedDelimiter\ceil{\lceil}{\rceil}
    \newcommand{\norm}[1]{\| #1 \|}
---

@@

On the other hand, several open questions about the use of embeddings within actuarial work remain, which we aim to 
address in this study. First, hyperparameter settings for embeddings, such as the dimensions of the embedding layer and
the use of regularization techniques such as dropout or normalization, that achieve optimal predictive performance has
not yet been studied in detail in the actuarial literature. In this work, we aim to study how embeddings using 
different settings perform in the context of a large-scale predictive modelling problem, and give guidance on the 
process that can be followed to determine this in other problems. Although neural network have been shown to achieve
excellent predictive accuracy on actuarial tasks, many actuaries still prefer to use GLM models for pricing tasks, 
thus, the issue of whether transferring embeddings to GLM models can achieve better performance is considered in this
paper. <!-- Traditional actuarial techniques such as credibility theory have been used to work with some types of categorical data, but no study has been -->
<!-- performed whether this can be applied within embeddings; here, we investigate whether performance is enhanced by applying credibility theory on embeddings 
relating to categorical variables with many labels. Whereas embedding layers are usually considered in the context of categorical data,the 
option exists to quantize numerical data and model it using embeddings, however, the results of doing so have not been investigated. -->
Finally, in the past several years, a new type of neural network architecture based on attention [@Vaswani2017] has
been successfully used on embeddings in the field of natural language processing and we incorporate attention 
based models into our predictive modelling example.

@@

During inference, or scoring, care must be taken to accommodate the categorical levels that are unseen
during training. While we allow for an extra key in the embedding dictionary, the initial
weights are not updated during training, which could lead to nonsensical predictions in the validation or test sets. To
circumvent this issue, before applying the trained model but after the final optimization epoch, we update the weights
for the unseen levels manually. In practice, a common approach is to take an average value 
of some sort of the weights of theother levels. In our case, we take the median (if there are multiple, the lower of
the two), but a mean, mean weighted by frequency of observations, or trimmed mean could be used.

In Table \@ref(tab:cv-results), we exhibit the cross-validated performance metrics of each of the models. We report
both the root mean squared error (RMSE) and the mean absolute error (MAE); we note that these are just a couple of many potential metrics to use, and are so chosen for their interpretability and prevalence in the literature for comparing
different types of models. Recall that, between the two metrics, RMSE penalizes large mispredictions than MAE. Also,
in the optimization process, GLM with a gamma distribution optimizes the deviance, rather than MSE directly.

@@

While linear regression (i.e., the specific case of GLM with a gaussian distribution and identity link function) is not
commonly used for severity modeling, we throw it into the mix as a benchmark, a practice we recommend in order to
check reasonableness of other models' results. Since it is possible for a linear
regression to output negative values, we cap the predictions below by 0.01 during scoring. Somewhat surprisingly, in 
the problem we
are considering, it performs similar to the more complex neural network models in the RMSE metric (which it optimizes
for, as the unit deviance of the normal distribution correspond to MSE.) Of course, this is not to discredit neural
networks or GLMs, as the experiments are performed without extensive hyperparameter tuning or feature engineering.

@@

```{r nn-3, echo = FALSE, fig.cap = "Architecture of Model 6.", out.width = "80%", fig.align = "center"}
knitr::include_graphics("images/tabt.pdf")
```

### Model 7: TabNet
The final model we consider in this subsection is the TabNet architecture. The structure of the model is quite similar to Model 6, except 
that a more specialized form of self-attention is used to select the most important covariates for the modelling problem.

@@

## Results

>>>>

---
title: Categorical Predictor Embeddings and Attention Based Neural Networks in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - categorical data
  - neural networks
  - embeddings
  - attention
  - severity modelling
  - National Flood Insurance Program 
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
header-includes: |
    \usepackage{amsmath}
    \usepackage{mathtools}
    \usepackage{bm}
    \usepackage{bbm}
    \usepackage{float}
    \DeclarePairedDelimiter\ceil{\lceil}{\rceil}
    \newcommand{\norm}[1]{\| #1 \|}
---

@@

On the other hand, several open questions about the use of embeddings within actuarial work remain, which we aim to 
address in this study. First, hyperparameter settings for embeddings, such as the dimensions of the embedding layer and
the use of regularization techniques such as dropout or normalization, that achieve optimal predictive performance has
not yet been studied in detail in the actuarial literature. In this work, we aim to study how embeddings using 
different settings perform in the context of a large-scale predictive modelling problem, and give guidance on the 
process that can be followed to determine this in other problems. Although neural network have been shown to achieve
excellent predictive accuracy on actuarial tasks, many actuaries still prefer to use GLM models for pricing tasks, 
thus, the issue of whether transferring embeddings to GLM models can achieve better performance is considered in this
paper. <!-- Traditional actuarial techniques such as credibility theory have been used to work with some types of categorical data, but no study has been -->
<!-- performed whether this can be applied within embeddings; here, we investigate whether performance is enhanced by applying credibility theory on embeddings 
relating to categorical variables with many labels. Whereas embedding layers are usually considered in the context of categorical data,the 
option exists to quantize numerical data and model it using embeddings, however, the results of doing so have not been investigated. -->
Finally, in the past several years, a new type of neural network architecture based on attention [@Vaswani2017] has
been successfully used on embeddings in the field of natural language processing and we incorporate attention 
based models into our predictive modelling example to demonstrate the use of these models for actuarial purposes.

@@

During inference, or scoring, care must be taken to accommodate the categorical levels that are unseen
during training. While we allow for an extra key in the embedding dictionary, the initial
weights are not updated during training, which could lead to nonsensical predictions in the validation or test sets. To
circumvent this issue, before applying the trained model but after the final optimization epoch, we update the weights
for the unseen levels manually. In practice, a common approach is to take an average value 
of some sort of the weights of the other levels. In our case, we take the median (if there are multiple, the lower of
the two), but a mean, mean weighted by frequency of observations, or trimmed mean could be used.

In Table \@ref(tab:cv-results), we exhibit the cross-validated performance metrics of each of the models. We report
both the root mean squared error (RMSE) and the mean absolute error (MAE); we note that these are just two of many 
potential metrics to use, and are so chosen for their interpretability and prevalence in the literature for comparing
different types of models. Recall that, between the two metrics, RMSE penalizes large mispredictions than MAE. Also,
in the optimization process, GLM with a gamma distribution optimizes the deviance, rather than MSE directly.

@@

While linear regression (i.e., the specific case of GLM with a gaussian distribution and identity link function) is not
commonly used for severity modeling, we add it into the mix as a benchmark, a practice we recommend in order to
check reasonableness of other models' results. Since it is possible for a linear
regression to output negative values, we cap the predictions below by 0.01 during scoring. Somewhat surprisingly, in 
the problem we
are considering, it performs similarly to the more complex neural network models in the RMSE metric (which it optimizes
for, as the unit deviance of the normal distribution correspond to MSE.) On the other hand, the performance of the linear model
is quite significantly worse on the MAE metric. Of course, this is not to discredit neural networks or GLMs, as the experiments 
are performed without extensive hyperparameter tuning or feature engineering.

@@

```{r nn-3, echo = FALSE, fig.cap = "Architecture of Model 6.", out.width = "80%", fig.align = "center"}
knitr::include_graphics("images/tabt.pdf")
```
### Model 7: TabNet
The final model we consider in this subsection is the TabNet architecture. The structure of the model is quite similar to Model 6, except 
that a more specialized form of self-attention is used to select the most important covariates for the modelling problem.

## Training and Inference {#training-inference-attn}
We follow a similar training setup for the three attention based models, to the neural networks discussed in the previous section. For Models 5 and 6, 
we utilize the \texttt{Adam} optimizer as before, however, due to the increased complexity of the models in this section,
we use learning rate of 0.001 (compared to 0.01 for the simpler models). Reflecting this reduced learning rate, 
we train on the analysis set for a maximum of 150 epochs, and select the best model based on the lowest loss achieved on the assessment. Moreover, since these models are
more complex than those tested in the previous section, we also add some minimal regularization using dropout [@srivastava2014dropout] to the attention layers 
and the last layer of the network. Dropout functions by setting some components of the network randomly to zero during each epoch of training. This forces each
component of the network to learn an independent representation of the data. In this section, we drop out components with a probability of 2.5\%. A final detail is that, 
for both of these models, we use an embedding dimension of size 16 for each of the categorical covariates, since the attention mechanism requires that the embeddings 
be of a common length. For the TabTransformer model, we use a row embedding of dimension 4.

For Model 7, we use the suggested training setup that utilizes the \texttt{Adam} optimizer with a learning rate of 0.02, and train for 15 epochs (since longer training
times did not lead to improved performance).

In Table \@ref(tab:cv-results-attn), we show the cross-validated performance metrics of each of the models, 
as well as the two best performing models from the previous section. As before, we report both the root mean squared error (RMSE) 
and the mean absolute error (MAE). The results show that Model 5, which simply adds attention to Model 4, achieves slightly worse
performance than Model 4 as measured by the RMSE, but using the MAE as a metric, outperforms Model 4. Thus, there is some room to argue that
for tabular data modelling that uses embeddings, the use of attention within models should be considered. Disappointingly, the TabNet 
architecture performs poorly, perhaps due to insufficient hyperparameter tuning, but indicating that ``out of the box", TabNet does not
appear to work well for this modelling task. Finally, the TabTransformer model shows the best performance of all models considered in 
this research, beating the others by a significant margin on both the RMSE and MAE metrics.

Table: (\#tab:cv-results-attn) Cross validation results for Models 5-7.

|Model                                                                          |   RMSE|    MAE|
|:-----------------------------------------------------------------------------|------:|------:|
|Model 4: MLP with multidimensional embeddings                                 | 60,973| 37,574|
|Model 5: Simple Attention                                                     | 61,022| 37,025|
|Model 6: TabNet                                                               | 62,938| 38,386|
|Model 7: TabTransformer                                                       | 59,900| 36,343|
|Linear regression, predictions capped below at 0.01                           | 60,879| 38,431|

In what follows, we focus on understanding the TabTransformer model in more detail.

@@
